---
title: "financial data analysis"
author: "Marie JULLIEN"
format: html
editor: visual
---

```{r}
#| message: false
library(here)
library(vroom) # or readr
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Data Loading
```{r}
#| message: false
clients <-vroom(here("data", "client.csv"))
accounts <-vroom(here("data", "account.csv"))
disp <-vroom(here("data", "disp.csv"))
```

## Joining data tables 
A collection of joining functions of the form `*_join`.

### Inner join 
```{r}
da <- inner_join(accounts, disp)
```
Look at the column that at the same name & operate the join over column that have the same name 

```{r}
cda <- inner_join(clients, da)
```

```{r}
cda <- inner_join(clients, da, by= join_by("client_id"), suffix=c("home", "bank"))
```

### Home district vs Bank district

```{r}
home_diff_bank <- cda |> filter (district_idhome!= district_idbank) |> nrow()
```
in the first join, all those people were excluded 

We observe that `r home_diff_bank` client have a bank in another district as the one they live in.

```{r}
cda |> 
  group_by(account_id)|> 
  distinct(district_idhome) |> 
  summarise(N=n()) |> 
  filter(N>1)
```
distinct -> see all the different value of the district home inside a group 

### Gender gap? 
```{r}
clients |>
  ggplot(aes(x=gender)) + geom_bar() + coord_flip()
```
=> no gender imbalance in the data set 
=> coord flix tourne le graph 

```{r}
cda |> 
  count(gender, type)
```
=> no gender effect bc distribution roughly the same 

```{r}
cda |> 
  count(gender, type) |>
  group_by(gender) |>
  mutate(freq=n/sum(n)) |>
  select(-n) |> 
  ungroup() |>
  pivot_wider(values_from= freq, names_from = type)
```
```{r}
cda |>
  ggplot(aes(y=gender, fill=type)) + geom_bar(position = "fill")
```

```{r}
chisq.test(cda|>pull(gender),
           cda |> pull(type))
```
H0 is not rejected 
